---
title: "Zoo classification"
author: "Satsawat N."
date: "January 03, 2017"
output:
  html_document: default
---

## Zoo Animal Classification {.tabset}
The aim is to use Machine Learning Methods to Correctly Classify Animals Based Upon Attributes.

### Overview

This data is available from [Kaggle](https://www.kaggle.com/uciml/zoo-animal-classification), which is published by [UCI ML](https://www.kaggle.com/uciml).


```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(caret)
library(gridExtra)
library(seriation)

rm(list=ls())
my.dir <- "D:/Projects/5.0 Personal project/19.0 Zoo classification/"
setwd(my.dir)

class.data <- read.csv("class.csv", header=T, sep=",", stringsAsFactors=F)
raw.data <- read.csv("zoo.csv", header=T, sep=",", stringsAsFactors=F)

colnames(raw.data) <- tolower(colnames(raw.data))
```

There are 2 datasets, namely: 

1. Class dataset - this provides the class name to the class type as well as the list of animals which are corresponding to the class.

```{r sample class, echo=FALSE}
head(class.data,2)
```

2. Zoo - this data consists of the animal names with their features and the class which they are belong. This is what we trying to predict.

```{r sample zoo, echo=FALSE}
head(raw.data, 2)
```

Note that we can use these 2 datasets to join each other to get full information.

```{r left join, echo=TRUE, eval=FALSE}
zoo.data <- merge(x = raw.data, y = class.data, by = "class_type", all.x = TRUE)
```

```{r sample join, echo=FALSE}
# Given class_type is what we will use, let's factor this variable
colnames(class.data) <- c("class_type", 
                          "no_of_animal_species_in_class", 
                          "class_name", 
                          "animal_name"
                          )

class.data <- select(class.data, 
                     class_type,
                     class_name)

raw.data$class_type <- as.factor(raw.data$class_type)
zoo.data <- merge(x = raw.data, y = class.data, by = "class_type", all.x = TRUE)
head(zoo.data, 3)
rm(raw.data, class.data)
```


### Visualize the data

We will visualize the provided features in the datasets for each class name. Note, we can join the class_name from class dataset to the zoo dataset. Below is the code snippet I used - you may need to manipulate the column names a bit.


```{r visual, echo=FALSE}

plotHist <- function(df, i) {
  data <- data.frame(x=df[[i]])
  p <- ggplot(data=data, aes(x=factor(x))) + 
    stat_count() + 
    xlab(colnames(df)[i]) + 
    ggtitle(colnames(df)[i]) +
    theme_bw() + 
    facet_wrap(~df[[19]]) +
    theme(axis.text.x = element_text(angle = 90, hjust =1))
  return (p)
}

lets.plot <- function(df, fun, ii, ncol=3) {
  pp <- list()
  for (i in ii) {
    p <- fun(df=df, i=i)
    pp <- c(pp, list(p))
  }
  do.call("grid.arrange", c(pp, ncol=ncol))
}

convert.to.numeric <- function(df, lst){
  for (i in lst){
    df[[i]] <- as.numeric(df[[i]])
  }
  return(df)
}

int.list <- names(zoo.data)[which(sapply(zoo.data, is.integer))]
zoo.data <- convert.to.numeric(zoo.data, int.list)

rm(int.list)

for(i in seq(3,18,1)) {
  lets.plot(zoo.data, fun=plotHist, ii=i, ncol=1)
}

```


### Supervised learning

Let's try to build supervised machine learning model. In this one, I will use **SVM** from [e1071](https://cran.r-project.org/web/packages/e1071/e1071.pdf) library to build the model.

Setting up the training (70%) and testing (30%) datasets. 

```{r train data prep, eval=FALSE}
intraining <- createDataPartition(y=zoo.data$class_type, p=0.7, list=F)
train.batch <- zoo.data[intraining,]
test.batch <- zoo.data[-intraining,]
```

Before going in the model part, due to the dataset size, I would like to do a quick check on each class sizing.

```{r quick check, echo=FALSE}
intraining <- createDataPartition(y=zoo.data$class_type, p=0.7, list=F)
train.batch <- zoo.data[intraining,]
test.batch <- zoo.data[-intraining,]
cat("----- Training batch ------")
table(train.batch$class_type)

cat("----- Testing batch -----")
table(test.batch$class_type)
```

Okay, this looks fine to me. Let's dive to the model set up part.

```{r svm, echo=FALSE}
library(e1071)
train.batch <- select(train.batch,
                      -animal_name,
                      -class_name)

train.x <- select(train.batch,
            -class_type)

train.y <- train.batch$class_type

set.seed(501)
```

```{r svm2}
svm.model <- svm(class_type ~ ., 
                 data=train.batch,
                 kernel = "radial")


summary(svm.model)

test.x <- select(test.batch,
            -class_type,
            -animal_name,
            -class_name)

test.y <- test.batch$class_type

test.pred <- predict(svm.model, test.x)
confusionMatrix(test.pred, test.y)
```



